The Adaptive Multi-scale Simulation Infrastructure - Design and Application

### 2016-01-25 doc/articles/amsi_biotissue/paper.tex
paper

% introduce scale-coupling topic to be discussed further in AMSI and methods sections
Numerically multi-scale simulations associate the computational implementations of single-scale physical models and cause them to interact by passing key physical terms between scales. 
This scale-coupling operation and may require the application of transformer functions to account for key physical differences between disparate scales. 
The primary purpose of such transformer functions is to convert physical coupling values produced by simulations operations at one scale to be meaningfully applied to simulation operations at the coupled, consuming scale. 
These transformer functions are typified by conversions between the unit metrics used by coupled scales. 
For example a micro-scale simulation will typically be implemented with units of $\mu$m while a macro-scale simulation might be implemented using units of cm.
Scale-coupling transformer functions may also be used for any additional coupling-term operations required to account for differences between coupled scales, such as dimensionalizing terms calculated in a dimensionless domain.

@@

A Scale is a data structure modelling the parallel state of a discrete simulation scale and all operations (or tasks) occurring at that scale. 
All Scales have a set of MPI\_Ranks associated with them through an intermediary structure called the RankSet. 
In the current implementation the RankSet of any two Scales in a simulation must be non-overlapping.  
A Scale is required for the specification of all other data structures in the AMSI libraries.

@@

@@

%% revised up to here, with sporadic edits after

@@

>>>>

% introduce scale-coupling topic to be discussed further in AMSI and methods sections
Numerically multi-scale simulations associate the computational implementations of single-scale physical models and cause them to interact by passing key physical terms between scales. 
This scale-coupling operation and may require the application of transformer functions to account for key physical differences between disparate scales. 
The primary purpose of such transformer functions is to convert physical coupling values produced by simulations operations at one scale to be meaningfully applied to simulation operations at the coupled, consuming scale. 
These transformer functions are typified by conversions between the unit metrics used by coupled scales. 
For example a micro-scale simulation will typically be implemented with units of $\mu$m while a macro-scale simulation might be implemented using units of cm.
Scale-coupling transformer functions may also be used for any additional coupling-term operations required to account for differences between coupled scales \cite{weinan2003heterogenous}, such as dimensionalizing terms calculated in a dimensionless domain \cite{stylianopoulos2007volume} \cite{chandran2007deterministic}.

@@

A Scale is a data structure modelling the parallel state of a discrete simulation scale and all operations (or tasks) occurring at that scale. 
All Scales have a set of MPI\_Ranks associated with them through an intermediary structure called the RankSet. 
In the current implementation the RankSet of all Scales in a simulation must be non-overlapping.  
A Scale is required for the specification of all other data structures in the AMSI libraries.

@@

Constructing a new multi-scale simulation from a set of legacy single-scale simulation codes involves several steps.
Each Scale in the simulation must have a 'main-like' function with the signature \verb|int (int *, char *[], MPI_Comm)| which initializes the various supporting libraries used by that single-scale code on the provided MPI\_Comm and then begins conducting the simulation at that Scale. 
Typically this function can simply be adapted from a main function already used by the single-scale simulation code, but initializing on the provided MPI\_Comm rather than MPI\_Comm\_World. 
Other MPI calls in the code operating on MPI\_COMM\_WORLD should be replaced with AMSI\_COMM\_LOCAL which is the current MPI\_Comm associated with the AMSI Scale associated with the single-scale code. 
Multi-scale interaction points are determined and as wells as calls to create the appropriate DataDistributions and CouplingPatterns, which may be created during initialization for simplicity and modified later during execution.
Calls to appropriately set DataDistribution values and modify CouplingPatterns should be added after adaptive procedures are conducted at the desired scale.
AMSI does not treat any Scale as controlling any other Scale, thus care must be taken to avoid two Scales conducting adaptive procedures on the same CouplingPattern at the same time, though both Scales may conduct such procedures as long as all operations from adaptive procedures at one Scale are completed prior to starting adaptive procedures at the second Scale. 
This allows the CouplingPattern to reach a globally valid state prior to undergoing additional modification. 
When a Scale has many scale-tasks operating it is necessary to write new code to manage the instantiation and management of the various scale-tasks, and update the relevant DataDistributions as required.

@@

@@

[major revisions end here, though there are sporadic edits herafter]
%% revised up to here, with sporadic edits after
